---
title: "topic #6, polysemous verbs with light and concrete senses."
author: "stschwarz"
date: "`r Sys.time()`"
output:
  revealjs::revealjs_presentation:
    self_contained: false
    css: reveal.css
    reveal_options:
      slideNumber: true
      previewLinks: true
    transition: none
bibliography: CORPUS-LX.bib

---

```{r setup-slide, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)

```

<!--<style>
.gdbar img {
  width: 300px !important;
  height: 150px !important;
  margin: 8px 8px;
 // color: fff;
  }-->
<!-- <style> -->
<!-- /*aside.gdbar { -->
<!--   background: white; -->
<!-- }*/ -->
<!-- </style> -->

# term paper draft
> "If onomasiological frequency measurements do indeed correlate with elicitation tests, 
potential impact would be immense. Researchers would be able to examine onomasiological frequencies in spoken corpora rather than performing elicitation tests. That possibility would facilitate cognitive research into languages and varieties around the world, without the necessity of in situ psycholinguistic testing, and would also encourage the creation of more spoken corpora." 
([@mehl_what_2021 p.23](https://doi.org/10.1515/cllt-2017-0039))

## method
### following corpus was used to determine frequencies:   

Q.1: Santa Barbara corpus of spoken American English, ([@ucsb_santa_2005](https://www.linguistics.ucsb.edu/research/santa-barbara-corpus))

the corpus data was downloaded from: [https://www.linguistics.ucsb.edu/sites/secure.lsit.ucsb.edu.ling.d7/files/sitefiles/research/SBC/SBCorpus.zip](https://www.linguistics.ucsb.edu/sites/secure.lsit.ucsb.edu.ling.d7/files/sitefiles/research/SBC/SBCorpus.zip)

### exploring scripts:
- [14015.concrete-abstract_HA.R](https://github.com/esteeschwarz/SPUND-LX/blob/main/corpusLX/14015-HA/14015.concrete-abstract_HA.R): creates corpus of downloaded .zip
- [sbc-postag.R](https://github.com/esteeschwarz/SPUND-LX/blob/main/corpusLX/14015-HA/sbc-postag.R): postags the corpus
- [get-freq-df.R](https://github.com/esteeschwarz/SPUND-LX/blob/main/corpusLX/14015-HA/get-freq-df.R): first evaluation
- [collexemes-01.R](https://github.com/esteeschwarz/SPUND-LX/blob/main/corpusLX/14015-HA/collexemes-01.R): 2nd evaluation script

```{r,echo=FALSE}
#load("data/trn.make.cpt.RData")
#load("data/light.ann.make.RData")
dtemp<-tempfile()
### B
download.file("https://github.com/esteeschwarz/SPUND-LX/raw/main/corpusLX/14015-HA/data/SBC.ann.df.RData",dtemp)
load(dtemp)
download.file("https://github.com/esteeschwarz/SPUND-LX/raw/main/corpusLX/14015-HA/data/SBC.20-sample.RData",dtemp)
load(dtemp)
# ifelse(exists("trndf.lm",-1),fetch<-"given",fetch<-"rerun")
# if(fetch=="rerun")
#    source("14015.concrete-abstract_HA.R")
download.file("https://github.com/esteeschwarz/SPUND-LX/raw/main/corpusLX/14015-HA/data/plotlist.RData",dtemp)
load(dtemp)
# load("data/plotlist.RData")
download.file("https://github.com/esteeschwarz/SPUND-LX/raw/main/corpusLX/14015-HA/data/collex.eval.RData",dtemp)
load(dtemp)
   

#load("~/boxHKW/21S/DH/local/SPUND/corpuslx/stefanowitsch/HA/data/trndf.lm.cpt.RData")
#load("data/SBC.20-sample.RData")
```

## corpus sample

```{r}
kable(smdf[c(4,11,1,2,6,5,19),])

```

## distribution analysis 
cf. [@mehl_what_2021 pp.11-14]

```{r,kable-01-dist-make,echo=FALSE,warning=F,}
######################################
kable(plotlist$dist)
#kable(kdf,caption = "distribution of lemma /make/ over corpora")
```

---

```{r fig-01-dist,echo=F,fig.cap="distribution of lemma /make/ over corpora. ICE data from study."}
# barplot(cbind(ICE.w=i.make.w,ICE.sp=i.make.s,SBC.sp=i.make.m),main="distribution: lemma /make/",legend.text = c("concrete use","light use"))
#plot.plots("dist")
#barplot(kdf,main="distribution: lemma /make/",legend.text = c("concrete use","light use"))
barplot(plotlist$dist,main="distribution: lemma /make/",legend.text = c("concrete use","light use"))
```

## semantic alternatives to make
- includes alternatives to /make/ that Mehl defined in his study 
- all occurences are counted 
  - Mehl contrasted only these alternate occurences which:

>"are defined as those verbs that occur in the corpus with the same concrete direct objects as make, (take, and give,) and with a roughly equivalent meaning." (cf. @mehl_what_2021 p.13)  

---


```{r}
kable(plotlist$trntable,caption = "semantic alternatives")

```

---

```{r fig-02-alt,echo=F, fig.cap="proportion of semantic alternatives to concrete /make/ in SBC"}
#trntable<-table(scb.ann.df$alt[scb.ann.df$light==0])
#par(las=3,cex=0.5,pin=c("1.5","1.5"))
#kable(trntable,caption = "semantic alternatives")
#barplot(trntable[c(1,2,3,4,5,6,7)]/sum(table(scb.ann.df$alt)),main = "SBC concrete /make/ vs. alternate",ylab = "% in corpus")
barplot(plotlist$trntable[c(1,2,3,4,5,6,7)]/sum(table(scb.ann.df$alt)),main = "SBC concrete /make/ vs. alternate",ylab = "% in corpus")
```

---

```{r fig-03-alt,echo=F, fig.cap="proportion of semantic alternatives, only equivalent meaning of alternates"}
#trntable<-table(scb.ann.df$alt[scb.ann.df$light==0])
#par(las=3,cex=0.5,pin=c("1.5","1.5"))
#kable(trntable,caption = "semantic alternatives")
#barplot(trntable[c(1,2,3,4,5,6,7)]/sum(table(scb.ann.df$alt)),main = "SBC concrete /make/ vs. alternate",ylab = "% in corpus")
barplot(plotlist$nouns.f.cp,main="semantic alternates w/ equivalent meaning",ylab = "% in corpus")
#plot.plots("alt.2")
```

---

```{r}
kable(plotlist$eval.semas[,1:9])

```

---

```{r fig-04-sema,echo=F, fig.cap="semasiological log.like of near synonymes to make"}
barplot(plotlist$eval.semas$COLL.STR.LOGL~plotlist$eval.semas$ns,xlab = "",ylab = "collex.log.like")
#plot.plots("alt.2")
```

## 2nd approach collexeme evaluation

```{r echo=F}
coll6.2<-collex.eval
#coll6.2[coll6.2$head_lemma=="make",]
#coll6.2[coll6.2$head_lemma=="make"&coll6.2$light==0,]
#coll6.2[coll6.2$head_lemma=="make",]
#coll6.2[coll6.2$head_lemma=="make"&coll6.2$light==1,]
#coll6.2[coll6.2$head_lemma=="make",]
make.array<-c("make","generate","produce","create","build")
eval1<-coll6.2[coll6.2$head_lemma%in%make.array&coll6.2$light==0,]
#coll6.2[coll6.2$head_lemma=="make",]
#coll6.2[coll6.2$head_lemma=="make"&coll6.2$light==0,]
eval1<-coll6.2[coll6.2$head_lemma%in%make.array&coll6.2$light==0,]
make<-sum(eval1$T[eval1$head_lemma=="make"])
build<-sum(eval1$T[eval1$head_lemma=="build"])
create<-sum(eval1$T[eval1$head_lemma=="create"])
produce<-sum(eval1$T[eval1$head_lemma=="produce"])
generate<-sum(eval1$T[eval1$head_lemma=="generate"])
#eval1
#sum(eval1$T[eval1$head_lemma=="make"])
#factor(plotdf$T)
#table(plotdf)
take.array<-c("take","bring","carry")
eval2<-coll6.2[coll6.2$head_lemma%in%take.array&coll6.2$light==0,]
take<-sum(eval2$T[eval2$head_lemma=="take"])
bring<-sum(eval2$T[eval2$head_lemma=="bring"])
carry<-sum(eval2$T[eval2$head_lemma=="carry"])
df<-factor(coll6.2$head_lemma)
#levels(df)
df<-length(levels(df))
#df<-177
#edge
#p_value_left = pt(q = -0.77, df = 15, lower.tail = TRUE)
get.p<-function(x)pt(x,df,lower.tail = F)
#get.p<-function(x)pt(x,df,lower.tail = T)
eval1$p<-unlist(lapply(eval1$T, get.p))
#eval1
eval2$p<-unlist(lapply(eval2$T, get.p))
#eval2
#x<-eval1$head_lemma[1]
sumobs<-function(x)sum(grepl(x,eval1[['head_lemma']]))
eval1$obs<-unlist(lapply(eval1$head_lemma, sumobs))
#obs1
#eval1
plotdf1<-data.frame(lemma=factor(eval1$head_lemma),p=eval1$p,obs=eval1$obs)
plotdf2<-data.frame(lemma=factor(eval2$head_lemma),p=eval2$p)
eval2$obs<-unlist(lapply(eval2$head_lemma, sumobs))

```

---

```{r fig-05-sema-T-make,echo=F, fig.cap="semasiological: sum T-score of near synonymes to make"}
barplot(cbind(make,produce,create,build,generate), main="absolute preference in concrete context",ylab="T score sum association strength")

```

---

```{r fig-06-sema-T-take,echo=F, fig.cap="semasiological: sum T-score of near synonymes to take"}
barplot(cbind(bring,carry,take),ylab="T score sum association strength", main="absolute preference in concrete context")

```

---

```{r fig-07-sema-p-make,echo=F, fig.cap="semasiological: p of near synonymes to make"}
par(las=3)
boxplot(plotdf1$p~plotdf1$lemma,varwidth=T,outline=F,xlab = "",ylab="p-value of collexeme association strength",main="binding of lemma /make/ in concrete noun context")

```

---

```{r fig-08-sema-p-take,echo=F, fig.cap="semasiological: p of near synonymes to take"}
#getwd()
par(las=3)
boxplot(plotdf2$p~plotdf2$lemma,varwidth=T,outline=F,xlab = "",ylab="p-value of collexeme association strength",main="binding of lemma /take/ in concrete noun context")

```

------

# B: references



