---
title: "14033.pub.essai"
author: "esteeschwarz"
date: "2024-01-08"
output: 
  md_document:
   variant: markdown_phpextra+backtick_code_blocks
   toc: true
   number_sections: true
   pandoc_args: ["--wrap=none"]
bibliography: comp.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# pub essai R to WP
publishing using R package *duncantl/RWordPress* [@lang_duncantlrwordpress_2023]. wks half: the pandoc citation features and footnotes which are generated by the *knit2wp()*-function of the *knitr* package [@yihui_knitr_2024] are not correctly converted to the post html.

## snc
1. knitXwp lib installed on mini.2:post wt lib

## workaround:
script to create html content of post:
```{r,eval=FALSE}

library(markdown)
library(rmarkdown)
pid<-0 # set postID of post created with knit2wp() or any existing post
rmd<-"sample.Rmd" # path of the R-Markdown file
wp.knit.post<-function(rmd,pid){
render(rmd) # renders the .Rmd according to the parameters specified in the YAML header of the .Rmd
md.ns<-gsub("\\.Rmd",".md",rmd)
p.content<-readLines(md.ns)
p.html<-mark(p.content) # renders the created .md file
editPost(pid,content = p.html) # actualises post content
}
```


yaml-header of the .Rmd rendered:
```{}
---
title: "14022"
author: "xxx"
date: "2024-01-08"
output: 
  md_document:
   variant: markdown_phpextra+backtick_code_blocks
   toc: true
   number_sections: true
   pandoc_args: ["--wrap=none"]
bibliography: sample.bib
---
```

## persistent solution
modified the content of the *knit2wp()* function as to use *render()* and *mark()* instead of *knit()* and *markdownToHTML()*. now works.

## R package
wrote small R package that does the same: <https://github.com/esteeschwarz/knitXwp>

## test image upload

```{r echo=F}
plot(cars)

```


---
## B: references