---
# seminararbeit im Seminar: Neuropragmatik, WS22/23 FUB
# Abgabe: 2022-11-20
title: "N-PRG conversation analysis"
subtitle: "Neuropragmatik WS22/23 FUB (Pulvermüller)"
author: "St. Schwarz"
date: "`r Sys.Date()`"
zotero: AVL_dyn
output: 
  html_document: 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    self_contained: no
    code_download: no
    css: "https://ada-sub.rotefadenbuecher.de/skool/public/essais/PRG01/style_HA.css"
    keep_md: yes
#bibliography: https://raw.githubusercontent.com/esteeschwarz/DH_essais/main/sections/DD/klemm_HA/top/klemm.bib
   # css: https://ada-sub.rotefadenbuecher.de/skool/public/papers/011/style_HA.css
#bibliography: "https://api.zotero.org/groups/4713246/collections/YTVL6QLM/items/top?format=bibtex"
bibliography: PRG_CA.bib
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval=TRUE, echo=FALSE, warning=FALSE,message=FALSE}
library(httr)
share <- runif(1)
x<-GET("https://api.zotero.org/groups/4713246/collections/YTVL6QLM/items/top?format=bibtex")
bib<-content(x,"text")
y<-tempfile("ref",fileext = ".bib")
writeLines(bib,y)


```
---
bibliography: "`r y`"
---

```{r eval=TRUE, echo=FALSE, warning=FALSE,message=FALSE}
#library(jsonlite)
library(syuzhet)
library(readr)
library(xfun)
library(httr)

root<-"~/boxHKW/21S/DH/"
#EWA,lapsi
root<-"~/boxHKW/UNI/21S/DH/"

expo<-paste0(root,"local/CA/PRG/expo/")
#setwd(local)
#src<-"wolf_FF_1.json"
#src<-paste0(local,"wolf_FF-LEN_1.json")
###
#mini
#setwd("~/boxHKW/21S/DH/")
#lapsi, ewa
#setwd("~/boxHKW/UNI/21S/DH/")

x<-GET("https://api.zotero.org/groups/4713246/collections/YTVL6QLM/items/top?format=bibtex")
bib<-content(x,"text")



```

## 1. head
the conversation is part of a 15 minutes planungsgespräch. the preliminary discussion revealed that one class intended to visit commune in the future (anschluszseminar) could be one in the section :german as foreign language: (DAF) that both participants considered relevant to their curriculum and in continuation of this classes (NPRG) program.

## 2. the transcript
the dialogue represents minutes 08:42 - 10:45 of the recording transcribed according to GAT2 conventions using EXMARALDA partitur editor [[@worner_exmaralda_2014]](https://exmaralda.org/de/) and ELAN [[@elan_elan_2022]](https://archive.mpi.nl/tla/elan).

#### 2.1 basic
``` {=html}

<iframe src="prime/CA_schnipsel_001.2.txt" style="border:2px solid black;" width="100%" height="400px"></iframe>
```

#### 2.2 analysis
``` {=html}

<iframe src="prime/CA_schnipsel_001.html" style="border:2px solid black;" width="100%" height="400px"></iframe>
```

## 3. some visualisations & statistics
### 3.1 TRP
Auswertung 1. der TRP (transition relevant places) im Dialogverlauf. Diese werden hier 2. nach Fourier-Transformation der Positionen abgebildet, so dasz die frequenzanalysierte (nach Fourier also relative Länge) der zwischen den TRP liegenden Dialogabschnitten sichtbar ist. 3. Die relativierte Verlaufkurve der Intonationsphrasenlänge; `rot = SP0, grün = SP1, blau = Differenzkurve`.

```{r eval=TRUE, echo=FALSE, warning=FALSE,message=FALSE}
#path1<-"local/CA/PRG/expo/"
src<-list.files(expo)
src<-paste0(expo,src)
trans_ns<-c("tier","type","speaker","text","start","end","category","label")
#dta<-read.csv2(src,header=F,sep = "\t",col.names = trans_ns)
dta <- read_delim(paste0(expo,"CA_schnipsel_001.tsv"), 
                  delim = "\t", escape_double = FALSE, 
                  col_names = trans_ns, trim_ws = TRUE)

###
trp<-dta$category=="TRP"
trp1<-dta$start[trp]
trp2<-c(dta$start[1],trp1)
trp3<-c(trp1,dta$end[length(dta$end)])
trp4<-trp3-trp2
barplot(trp4, main = "TRP evaluation",ylab = "absolute length of phase",xlab="flow of conversation")
library(syuzhet)
trp5<-get_transformed_values(trp4)
#plot(trp5,type = "h")
plot(trp5,type="h",main="TRP evaluation",ylab="relative length of phase",xlab="index of conversation",col=2)
points(trp5, cex = .2, col = "dark red")
###wks.
spk0<-dta$speaker=="SPK_SP0"&dta$category=="main"
spk1<-dta$speaker=="SPK_SP1"&dta$category=="main"
spk2<-dta$end[spk0]
spk3<-dta$end[spk1]
spk4<-dta$start[spk0]
spk5<-spk2-spk4 #
speechA0<-sum(spk0)
speechA1<-sum(spk1)

#spk5[spk0]
# cat("speech acts M: ",speechA0,"\n")
# cat("speech acts F: ",speechA1,"\n")
#sum(spk1)
spk6<-dta$start[spk1]
spk7<-spk3-spk6
sumSPK1<-sum(spk7)
sumSPK0<-sum(spk5)
# cat("absolute time speech acts M: ",sumSPK0,"\n")
# cat("absolute time speech acts F: ",sumSPK1,"\n")

spk8<-get_transformed_values(spk5)
spk9<-get_transformed_values(spk7)
plot(spk8,type="h",main="utterance length ~ dialog",ylab="relative length of IP (ms)",xlab="index of conversation",col=2)
par(new=T)
plot(spk9,type="h",ann = F,col=3,yaxt="n")
spk10<-(1/spk9-1/spk8)
#spk10<-sqrt(spk10*spk10)
#spk10<-get_transformed_values(spk10)
par(new=T)
plot(spk10,type="l",ann = F,col=4,yaxt="n")
```

Fourier transformation of data cf. "syuzhet" R-Package: [[@jockers_revealing_2015]](https://www.matthewjockers.net/2015/02/02/syuzhet/)

### 3.2 some numbers
```{r echo = FALSE,warning=FALSE}
###semantic statistics
library(stringi)
tsp0<-dta$text[spk0]
tsp01<-unlist(stri_split(tsp0,regex = " "))
regx1<-"[^A-Za-zäöü]"
tsp02<-gsub(regx1,NA,tsp01)
tna<-!is.na(tsp02)
words_spk0<-sum(tna)

tsp0<-dta$text[spk1]
tsp01<-unlist(stri_split(tsp0,regex = " "))
regx1<-"[^A-Za-zäöü]"
tsp02<-gsub(regx1,NA,tsp01)
tna<-!is.na(tsp02)
words_spk1<-sum(tna)

#length(tsp01)
speval<-data.frame()
speval<-cbind("speechacts"=c(speechA0,speechA1),"speech time (sec) "=c(round(sumSPK0),round(sumSPK1)),"words"=c(words_spk0,words_spk1))
rownames(speval)<-c("SP0","SP1")
print(speval)
# xfun::pkg_load2(c("htmltools", "mime"))
# xfun::embed_files(c('N-PRG_CA_001.Rmd',"prime/CA_schnipsel_001.tsv","prime/CA_schnipsel_001.2.exb","prime/CA_schnipsel_001.2.txt","prime/CA_schnipsel_001.2.rtf"))

```
---
## A. notes
## B. REF:
<div id="refs"></div>

## C. annex
transcript files, tables & source code:
```{r echo = FALSE,warning=FALSE}
xfun::pkg_load2(c("htmltools", "mime"))
xfun::embed_files(c('N-PRG_CA_001.Rmd',"prime/CA_schnipsel_001.tsv","prime/CA_schnipsel_001.2.exb","prime/CA_schnipsel_001.2.txt","prime/CA_schnipsel_001.2.rtf"))
```

```{=html}
<!--			<p id="zt_content">temp</p>

<script>  
//function import_zt(block_id,api){
    var xhttp = new XMLHttpRequest();
      xhttp.open("GET", 'https://api.zotero.org/groups/4713246/collections/YTVL6QLM/items/top?format=bib&linkwrap=1', false);
      xhttp.send();
      document.getElementById("zt_content").innerHTML = xhttp.responseText;
    //  document.getElementById("show_button").style.display="none";
      //document.getElementById("hide_button").style.display="inline";
  //  }
</script>
-->
```

